MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites \\ Supplementary Material

### 2014-12-11 Rebuttal.tex

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{images_pvis/Vesselness}
\caption{Left: Xray image of vasculature, middle-left: calculated vesselness of the left image, middle-right: calculated vesselness after inversion of the grey-scale map. Right image after subtracting reference from left image.}
\label{fig:Vesselness}
\end{figure}

\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
	\includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
	\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
	\label{fig:data-char-rebuttal}
\end{figure}

\color{red}
Q2. In the examples presented, it seems only one type of composite was studied. For instance, there is no example showing a composite with fiber bundles that cross each other at an angle $<$ 90 degrees. When two bundles are almost parallel, will the current algorithm be sufficient to separate them apart? The authors are encouraged to provide more than one fiber configurations.
\color{black}

Answer. Due to the manufacturing process, a vast majority of composites have bundles crossing at ~90 degree angles. With the separation factors being width, spacings and compactness of fiber bundles. These variations are well represented in our data sets. We maintain that the technique is useful and applicable to a large percentage of composite polymers. We have demonstrated the technique on three datasets, two of them are in the paper and the third is in the supplemental section due space  constraints. Getting additional datasets at this time is not possible

In dataset 1, Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 shows a common example of variations in the datasets which we extract accurately. 
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{images_pvis/clustering.pdf}
\caption{Fiber bundles extraction}
\label{fig:angle_bundles}
\end{figure}

@@

\color{red}
  It seems to me that the extracted fibers may not be real fibers. Is this
  true? If so, how well can one trust the weaving patterns extracted based
  on the imaginary fibers? 
  \color{black}
  Answer. 
Yes, the extracted fibers cannot be mapped to real fibers. One of the goals  we  focused on was to move away from extraction of individual fibers, instead work on direct reconstruction of bundles. The design decisions adheres to extracting reliable bundles directly from the images. 
 Developing methods which can precisely compute the deviations from original bundles remains a future direction of the project.

\color{red}
The authors claim that their method can deal with low resolution data.
But this might be in the eyes of the beholders, since in my opinion the example data in the paper all have fairly high resolutions. Perhaps a comparison with a really high resolution data using this method and other existing methods could diffuse the doubt.
\color{black}

Answer. Figure~\ref{fig:glass_fibers} shows an example where each fiber is clearly demarcated compared to the type of data we address in the paper (Figure~\ref{fig:data-char-rebuttal}).
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{images_pvis/glass_fibers}
\caption{High Resolution fiber}
\label{fig:glass_fibers}
\end{figure}

\begin{figure}[htp]
\centering
 \includegraphics[width=0.4\linewidth]{imagesMT2014/glass-cluster-a.eps}
  \includegraphics[width=0.4\linewidth]{imagesMT2014/glass-cluster-b.eps}
   \includegraphics[width=0.4\linewidth]{imagesMT2014/glass-scalar.eps}
  \caption{ Fig  A and C shows the results fiber extraction. B and D shows the results of clustering with clipping planes.}\label{fig:closeup-full}
\end{figure}

### 2014-12-11 chapters/MetaTracts.tex

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx} 
A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Basser2000, Mori1999, Mori2002}. Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion. 
This is usually done by solving a Runge-Kutta integration (typically second or fourth order). The principal diffusion direction at each discrete location is interpolated to form a continuous velocity field. 
Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2002,Basser2000} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data. 
Mori et al.~\cite{Mori1999,Mori2002} divided reconstruction techniques into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and \removed{choosing} a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers \replacedWith{that}{which} are close for certain distance and then diverge, are clustered separately. Brun et al.~\cite{Brun2004} use\removed{s} normalized cuts along with a pairwise distances measure computed using a 9\removed{-}D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{Brun2004,jonasson2005, ODonnell2007}.
Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique, Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}), can be applied to the problem of organizing fiber tracts data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber tracts. In the ideal case fiber tracts belonging to the same bundle must remain ``close" to each other in the lower dimensional space. Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of point wise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}. 
The agglomerative hierarchical clustering method~\cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results. 
There are a number of difficulties in hierarchical clustering. First, computing all pairs\added{'} distances for tracts to generate the distance matrix is time consuming~\cite{Garyfallidis2012}. Second, a ``correct'' distance measure to compare tracts must be chosen. Third, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

@@

\subsection {Visual Analysis and Modeling of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores and deals with fiber extraction from high resolution data where the individual fibers are clearly discernible. Fritz et al.~\cite{Fritz2009} proposed \removed{interactive} workflows for \replacedWith{non destructive testing}{NDT} practitioners to explore and quantify steel fibers in reinforced sprayed concrete. This approach allows analyzing fiber orientations based on direction transfer functions. Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations. Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Lomov et al.~\cite{Lomov2010, Lomov2001} and Straumit et al.~\cite{Straumit2014} \removed{in recent papers}, discussed problems and current available solutions in geometric modeling of three dimensional composites. Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position. The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions. 
%Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research. Recently, similar techniques have been applied to fiber tracing in composite materials. 
To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of \replacedWith{composite fibers, which is the main scope of this work}{CFRP laminates}.

@@

The input to the first stage of our pipeline is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location \replacedWith{becomes}{gets} associated with an unit vector representing the local orientation and a real value [0,1] representing a measure of the reliability of the local orientation. We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.
\replacedWith
{The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}
{The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}

\replacedWith
{
Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute a similar metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  
}
{
Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the \textit{vesselness} criterion. In order to determine \textit{reliable Hessians} for our approach, we make use of the vesselness criterion as follows: Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.
}

\begin{equation}\label{RA}
\mathcal{R_{A}}=\frac{\textrm{Largest  Cross Section}\big/ \pi}{{\textrm{Largest Axis Semi-length}}^{2}}=\frac{|\lambda_{2}|}{|\lambda_{3}|}
\end{equation}

@@

Grid locations where the $R_{H}$ is above a cutoff threshold are marked as regions with reliable orientations (see Sec.~\ref{sec:param_choices} on parameter choices).

@@

In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold. The unit vector representing the principal direction has been mapped to RGB color space. 
%Figure~\ref{fig:reliable_hessian}a shows the entire data set. 
Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue. Figure~\ref{fig:reliable_hessian}b and c show 2D slices along the Z and X axes respectively. Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest. Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and therefore considered as unreliable. The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\includegraphics[width=\linewidth]{images_pvis/reliable_hessian.pdf}
 \vspace{-1.5em}
\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b\added{,c}) 2D slice\added{s} along Z\replacedWith{ and (c) along}{-} and X axis. (d) \replacedWith{Magnified}{Zoom in of the green} region marked in \replacedWith{green square}{(c)}.}
\label{fig:reliable_hessian}
\vskip-0.2cm
\end{figure}

@@

\subsection {Orientation based clustering}
%We use the inherent features expressed as weaving pattern of our data to pre-cluster the metaTracts into the number of classes based on the major directions of the pattern.
%It is important to note that this step broadly divides the MetaTracts into classes based on the major orientations and not individual fiber bundles. 
This step divides the individual MetaTracts into classes based on their major orientations. In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps \replacedWith{which was}{as} originally introduced by Belkin and Niyogi \cite{Belkin01}. An eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
Let $G$ be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem $L\textit{\textbf{f}}=\lambda D\textit{\textbf{f}}$,
%\begin{equation}\label{equn:eigenMaps}
%L\textbf{f}=\lambda D\textbf{f}
%\end{equation}
where $D$ is the diagonal weight matrix and $L$ is the Laplacian matrix. The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space (see Sec.~\ref{sec:param_choices} for values of m).
In our case, each MetaTract is a data point. We introduce a simple \textit{orientation based similarity measure}. There is no spatial information involved, just a partition of the MetaTracts based on orientation which is already inherent in the data.
Following this, given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts. The edge weights give a distance matrix representing the distance between each pair of points. Using the Belkin and Niyogi algorithm we ``embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix. 
%
%August 13 2014
%For all our tests the fibers were then embedded in a $\eta$ low dimensional space. The number of dimensions in the high dimensional space is equal to the number of datapoints or the number of MetaTracts in our case.
%For all our test cases we set $\eta$ to be five (see sec~\ref{sec:param_choices} for parameter choices for $\eta$.).
%  
We used conventional K-means for clustering this lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
For our test case there are two major directions of the fiber bundles and $K$ was set to 2. Due to the advantages offered by the dimensionality reduction, even if there are curved fiber bundles the user has to provide just the major fiber bundle directions of the weaving pattern. Figrue.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes. As is expected there is a clear distinction based on fiber orientation. Figure~\ref{fig:orientation_clustering}a,c shows the MetaTracts colored according to clustering results.

@@

\subsection{Choice of clustering techniques}
\label{subsec:clus_choice}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering. As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large  incorrect clusters and small(low-cardinality) outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top fiber bundles incorrectly cluster together.
%
Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step into 10, 15, and 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together. Even at $h=20$, no correct bundles are identified.
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only (a)hierarchical and (b)\added{dimensionality reduction followed by} K-means clustering methods to MetaTracts for various numbers of clusters \added{(distance measure is minimal directed Hausdorffs.}}
	\label{fig:comparison}
	\vskip-0.2cm
\end{figure}

>>>>

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx} 
A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Mori1999, Mori2002}. Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion. 
This is usually done by solving a Runge-Kutta integration (typically second or fourth order). The principal diffusion direction at each discrete location is interpolated to form a continuous velocity field. 
Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2002} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data. 
Mori et al.~\cite{Mori1999,Mori2002} divided reconstruction techniques into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, similarity measures such as proximity between fibers are used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and \removed{choosing} a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers \replacedWith{that}{which} are close for certain distance and then diverge, are clustered separately. Brun et al.~\cite{Brun2004} use\removed{s} normalized cuts along with a pairwise distances measure computed using a 9\removed{-}D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{Brun2004,jonasson2005, ODonnell2007}.
Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique, Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}), can be applied to the problem of organizing fiber tracts data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber tracts. In the ideal case fiber tracts belonging to the same bundle must remain ``close" to each other in the lower dimensional space. Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of point wise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}. 
The agglomerative hierarchical clustering method~\cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results. 
There are a number of difficulties in hierarchical clustering. First, computing all pairs\added{'} distances for tracts to generate the distance matrix is time consuming~\cite{Garyfallidis2012}. Second, a ``correct'' distance measure to compare tracts must be chosen. Third, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

@@

\subsection {Visual Analysis and Modeling of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fiber extraction from high resolution data where the individual fibers are clearly discernible. Fritz et al.~\cite{Fritz2009} proposed interactive workflows%for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. %This approach allows analyzing fiber orientations based on direction transfer functions. 
Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites.%They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Lomov et al.~\cite{Lomov2010} discusses the problems and current available solutions in geometric modeling of three dimensional composites.
%Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position? The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.

%Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research. Recently, similar techniques have been applied to fiber tracing in composite materials. 
%To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers.

@@

The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location \replacedWith{becomes}{gets} associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation. We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.
\replacedWith
{The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}
{The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}

The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute 
\color{red}{ the same metric} \color{blue}{a similar metric}\color{black}. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  
\begin{equation}\label{RA}
\mathcal{R_{A}}=\frac{\textrm{Largest  Cross Section}\big/ \pi}{{\textrm{Largest Axis Semi-length}}^{2}}=\frac{|\lambda_{2}|}{|\lambda_{3}|}

@@

Grid locations where the $R_{H}$ is above a cutoff threshold are marked as regions with reliable orientations (see Sec.~\ref{sec:param_choices}).

@@

In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold. The unit vector representing the principal direction has been mapped to RGB color space. Figure~\ref{fig:reliable_hessian}a shows the entire data set. %Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue. 
Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively. Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest. Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable. The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\includegraphics[width=\linewidth]{images_pvis/reliable_hessian.pdf}
 \vspace{-1.5em}
\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b\added{,c}) 2D slice\added{s} along Z\replacedWith{ and (c) along}{-} and X axis. (d) \replacedWith{Magnified}{Zoom in of the green} region marked in \replacedWith{green square}{(c)}.}
\label{fig:reliable_hessian}
\vskip-0.2cm
\end{figure}

@@

\subsection {Orientation based clustering}
%We use the inherent features expressed as weaving pattern of our data to pre-cluster the metaTracts into the number of classes based on the major directions of the pattern.
%It is important to note that this step broadly divides the MetaTracts into classes based on the major orientations and not individual fiber bundles. 
This step divides the individual MetaTracts into classes based on their major orientations. In order to cluster MetaTracts going in the same directions, we use a spectral embedding technique called Laplacian eigenmaps \replacedWith{which was}{as} originally introduced by Belkin and Niyogi \cite{Belkin01}. An eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
Let $G$ be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem $L\textit{\textbf{f}}=\lambda D\textit{\textbf{f}}$,
%\begin{equation}\label{equn:eigenMaps}
%L\textbf{f}=\lambda D\textbf{f}
%\end{equation}
where $D$ is the diagonal weight matrix and $L$ is the Laplacian matrix. The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an $m$-dimensional space (see Sec.~\ref{sec:param_choices} for values of m).
 In our case, each MetaTract is a data point. We introduce a simple \textit{orientation based similarity measure} to partition the  MetaTracts based on orientation already inherent in the data.
Given a pair of  MetaTracts, we define the edge weights between two tracts as the cosine of the maximum angle between the local orientations ($N_P$) of all pairs of start points ($C_P$) between the two MetaTracts. The edge weights give a distance matrix representing the distance between each pair of points. Using the Belkin and Niyogi algorithm we ``embed" these points in a low dimensional space where the Euclidean distance between points approximates the distance between points given by the distance matrix. 
We used conventional K-means for clustering this lower dimensional space. Here $K$ is supplied by domain knowledge of the number of major fiber bundle directions of the woven structure.
For our test case there are two major directions of the fiber bundles ($K=2$). Due to the advantages offered by the dimensionality reduction, even if there are curved fiber bundles the user has to provide just the major fiber bundle directions of the weaving pattern. Figure.~\ref{fig:orientation_clustering}b shows the result of the K-means clustering with the data points projected to the top three eigenvectors as the major axes. As is expected there is a clear distinction based on fiber orientation. Figure~\ref{fig:orientation_clustering}a,c shows the MetaTracts colored according to clustering results.

@@

\subsection{Choice of clustering techniques}
\label{subsec:clus_choice}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters without first performing orientation clustering (ground truth is 11 clusters). As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large  incorrect clusters and small (low-cardinality) outliers. When number of clusters ($h$) is 10, two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top fiber bundles incorrectly cluster together.
%
Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering by proximity alone, after being embedded in a $m$-dimensional (lower) space, without a second hierarchical clustering step into 10, 15, and 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together. Even at $h=20$, no correct bundles are identified.
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only (a)hierarchical and (b)\added{dimensionality reduction followed by} K-means clustering methods to MetaTracts for various numbers of clusters \added{(distance measure is minimal directed Hausdorffs.}}
	\label{fig:comparison}
	\vskip-0.2cm
\end{figure}
